---
title: "Use linear regression to predict condominum price in NYC"
output: html_notebook
---
This project uses linear regression to predict the condominium prices in NYC.
And compare the regressions across different boroughs.

load the libraries

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(broom)
library(purrr)
```

load the data
select only the condos
```{r}
NYC_property_sales <- read_csv("NYC_property_sales.csv")
NYC_condos <- NYC_property_sales %>%
  filter(building_class_at_time_of_sale == "R4") %>%
  filter(!is.na(sale_price)) %>%
  filter(!is.na(gross_square_feet))

```
use summary and boxplot to exam the data

Manatan has two outliers that skewed the entire regression line
Their sales price (>= 200 million) and gross_square_feet were both outliers
1. 165 East 66th Street represents the sale of many units (gross_square_feet = 233578)
2. 220 Central Park South, 50 (gross_square_feet = 23029)

Investigate the outliers in Bronx, Brooklyn to see whether they represent multi-units. None of those sales represent multi-units, so they were kept.

Investigate duplicate records. No duplicate records
multi-unit sales called multi_unit_sales that contains sale records where there is three-or-more records when grouped by sale_price and sale_date

```{r}
summary(NYC_condos$gross_square_feet)

ggplot(data = NYC_condos, aes(x = borough, y = gross_square_feet)) + 
  geom_boxplot()

NYC_condos %>%
  group_by(borough) %>%
  summarize(total = n())


NYC_condos %>%
  filter(borough == "Manhattan") %>%
  arrange(desc(sale_price)) %>%
  head()

NYC_condos <- NYC_condos %>%
  filter(sale_price < 200*(10^6))

NYC_condos %>%
  filter(borough == "Bronx") %>%
  arrange(desc(sale_price)) %>%
  head()

NYC_condos %>%
  filter(borough == "Brooklyn") %>%
  arrange(desc(sale_price)) %>%
  head()

multi_unit_sales <- NYC_condos %>%
  group_by(sale_price, sale_date) %>%
  filter(n()>=3) %>%
  arrange(desc(sale_price))

NYC_condos2 <- NYC_condos %>%
  group_by(sale_price, sale_date) %>%
  filter(n()<3) %>%
  arrange(desc(sale_price))
  
```
scatter plot
use different xlim to examine the pattern in different range

Observations: the direction between gross_square_feet and sale_price is positive in all boroughs
The strengths are strong in staten Brons, Manhattan, island, weak in Brooklyn and Queens


```{r}
ggplot(data = NYC_condos, aes(x = gross_square_feet, y = sale_price, color = borough)) + 
  geom_point() + 
  scale_y_continuous(labels = scales::comma)+
  geom_smooth(method= "lm", se = FALSE)+
  facet_wrap(~borough,scales = "free")
```
build a linear model to use gross_square_feet to predict sale_price
Examine whether removing the multi_unit_sales would make the prediction better or worse

Observation:
The original NYC condo has 149 record of multi unit sale records, in which the units were sold at the same
exact price on the same day, even though each unit has different gross square feet. When we are building a linear regression model to use gross square feet to predict the sale price of the condos, we considered both scenaria, 1, using all the NYC condos with the multi unit sale records included, and we call it the original data, and 2, use the data without the multi unit sale recrods, and we'll call it the trimmed data.

Comparing the two models from both the original and the trimmed data, both models indicated that the gross square feet is an significant predictor to the sale price. But the trimmed data produced a more accurate model than the original data, supported by the following comparisons:

1. The standard errors of the trimmed versus original models are 34.23 vs. 47.42. A smaller standard error means a smaller difference between the predicted value and the observed value
2. The confidence intervals of the original model are 3786 to 3972. It means for every square foot increase, the sale price will mostly likely to increase in the range of $3886 to $3972. with a range of 186. The confidence intervals of the trimmed data are $3972 to $3796, and the width of the range is 134. A smaller range means a more accurate prediction.

```{r}
NYC_condos_lm_original <- lm(sale_price ~ gross_square_feet, data = NYC_condos)
summary(NYC_condos_lm_original)
confint(NYC_condos_lm_original)[2,]

NYC_condos_lm <- lm(sale_price ~ gross_square_feet, data = NYC_condos2)
summary(NYC_condos_lm)
confint(NYC_condos_lm)[2,]

```

replot using the data after the multi-unit-sale have been removed
Observation, the regression line in Brookly was improved. 
Conclusion, since the multi unit sale records are composed of units that were sold at the same price on the same day, but with different number of gross square feet, on its own there were any relationship between sale price and gross square feet. Thus removing these record improvied the linear regression of using gross square feet to predict sale price

```{r}
ggplot(data = NYC_condos2, aes(x = gross_square_feet, y = sale_price, color = borough)) + 
  geom_point() + 
  scale_y_continuous(labels = scales::comma)+
  geom_smooth(method= "lm", se = FALSE)+
  facet_wrap(~borough,scales = "free")
```

generate models for each boroughs separately.
Conclusion: all models are significant, indicating that gross square feet can predict sale price
The predictability is the strongest in Mahattan, in which the gross square feet can explain 63% of the total price variance

The predictablity decreases in the order of Bronx (59%), Brooklyn (52%), Staten Island(49%), Queens(34%)

The value of the slope indicate that for every square foot increase in size, how much the price would go up. Comparing the five boroughs, the price increase per square foot decreases in the order of: Manattan ($4054), Brooklyn($1285), Queens($732), Bronx($649),Staten Island($289).
```{r}
NYC_condo2_by_borough <- NYC_condos2 %>%
  group_by(borough) %>%
  nest()%>% 
  mutate(linear_model = map(.x = data, 
                            .f = ~lm(sale_price ~ gross_square_feet, 
                            data = .)))%>% 
  mutate(tidy_coefficients = map(.x = linear_model, 
                              .f = tidy, 
                              conf.int = TRUE))%>% 
  mutate(tidy_summary_stats = map(.x = linear_model,
                                  .f = glance)) %>% 
  mutate(data_augmented = map2(.x = linear_model, 
                          .y = data, 
                          .f = augment))

nyc_tidy_coefficients <- NYC_condo2_by_borough %>%
unnest(cols = tidy_coefficients)

nyc_summary_stats <- NYC_condo2_by_borough %>%
unnest(cols = tidy_summary_stats)

nyc_augmented <- NYC_condo2_by_borough %>%
select(borough, data_augmented) %>%
unnest(cols = data_augmented)

```
